---
title: "Baseline ABCD v5.1 prep"
author: "Eileen Xu"
date: "2024-09-17"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, results = "asis", warning = F)
library(tidyverse)
library(kableExtra)
options(knitr.kable.NA = '')
library(summarytools)
```

## Preparing baseline ABCD data for elastic net models

```{r}
dat = readRDS("G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/predictors_unrelatedIDs.rds")
baseline = dat %>% filter(eventname=="baseline_year_1_arm_1") %>% select_if(~ !all(is.na(.))) |> droplevels()
```

Total N = 9802 \
10% missing would be N = 980 missing \
20% = 1960 \
30% = 2940 \
40% = 3920 \
50% = 4901 \

Due to TLFB screening out those who haven't heard of or tried the drug (more than a sip or a puff), NAs most likely to be 0. With how rare these events are at this age though (`r round(min(baseline$interview_age, na.rm=T)/12, 2)` - `r round(max(baseline$interview_age, na.rm=T)/12, 2)`) years, it may be worth removing this variable and examining it at year 2 instead of baseline. Similarly, it may be better to remove illicit drugs from baseline analyses. When restricting variables to have no more than 20% missing data, tlfb for alcohol, cannabis and number of use days are excluded as predictors. Tobacco puff is still included.

Note on the "gender" variable: this is coded based on child's (parent-) reported gender identity. M includes cis and trans males, F includes cis and trans females, GNC = gender non-conforming, combines "GNC" and "Different" from ABCD. I will include both sex at birth ('birthsex')  and gender ('gender') in the saved data but it would be nice to use gender (might have to remove the few GNC as the interactions wouldn't really make sense with so few) - can run with birthsex as a sensitivity analysis. 

```{r}
## remove any variables with >20% missing data?
baseline = baseline |> select_if(~ sum(is.na(.))<0.2*nrow(baseline))

baseline = baseline %>% select(-c(eventname, illicit, site_id_l, rel_family_id, rel_birth_id, interview_date, visit_type, gender_id))

names(baseline) = c("src_subject_id", "tobacco_puff", "weightcontrol_ksads",
                    "witness_comm_violence", "death_loved_one",
                    "witness_dv", "s_abuse", "p_abuse", "emot_abuse",
                    "serious_accident", "sleep_hrs", "bmi", "needed_food",
                    "income", "parent_ed", "birthsex", "gender",
                    "race_ethnicity","area_depriv", "comm_safety",
                    "days_active", "agemths", "cbcl_internalising",
                    "cbcl_dsm5_depress")

# remove empty levels
baseline = droplevels(baseline)

# remove cbcl missing data
baseline = baseline %>% filter(!is.na(cbcl_internalising))

sapply(baseline, function(x) sum(is.na(x))) |> kbl(col.names = c("", "N missing")) |> kable_styling() |> scroll_box()
```

Little's MCAR test for continuous variables:

```{r}
baseline |> select_if(~ is.numeric(.)) |> naniar::mcar_test() |> kbl() |> kable_styling()
```

I plan on keeping the (5-level) ethnicity variable in for mice imputation, but will remove it from the final model. I think that I should prioritise analysis in the complete sample though? The imputed missing data is influenced by the other variables in the data (and thus entered into the model) and when it comes to longitudinal analysis it might make things overly complicated. But it's useful to keep the mice code here anyway for now at least, though it might get removed later.

```{r}
## split into train and test samples ##
index = caret::createDataPartition(baseline$cbcl_internalising[which(is.na(baseline$cbcl_internalising)==F)], 
                            p = 0.2, list = F, times = 1)
traindat = baseline[-index,]
testdat = baseline[index,]
```

## Descriptive statistics

### Full sample

```{r}
baseline[,-1] |> dfSummary(plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```

### Training data:

```{r}
traindat[,-1] |> summarytools::dfSummary(plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```

### Test data:

```{r}
testdat[,-1] |> summarytools::dfSummary(plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```


```{r eval=FALSE}
saveRDS(traindat, "G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/baseline_traindat.rds")
saveRDS(testdat, "G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/baseline_testdat.rds")
```

