---
title: "Y2 ABCD v5.1 prep"
author: "Eileen Xu"
date: "2024-09-17"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, results = "asis", warning = F)
library(tidyverse)
library(kableExtra)
options(knitr.kable.NA = '')
library(summarytools)
```

## Preparing year 2 ABCD data for elastic net models

```{r}
dat = readRDS("G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/predictors_unrelatedIDs.rds")
y2 = dat %>% filter(eventname=="2_year_follow_up_y_arm_1") %>% select_if(~ !all(is.na(.))) |> droplevels()
```

Total N = `r nrow(y2)` \
10% missing = `r nrow(y2)*0.1` \
20% missing = `r nrow(y2)*0.2` \
30% missing = `r nrow(y2)*0.3` \
40% missing = `r nrow(y2)*0.4` \
50% missing = `r nrow(y2)*0.5` 

```{r}
y2 |> select_if(~ is.numeric(.)) |> summary(.) |> kbl() |> kable_styling()
```

Risk factors with >20% missing data were removed.

TLFB alcohol includes data for `r sum(!is.na(y2$tlfb_cal_scr_alc_ud))` individuals, TLFB cannabis includes data for `r sum(!is.na(y2$su_tlfb_cal_scr_mj_days_yr))`. Total use days has data for `r sum(!is.na(y2$tlfb_cal_scr_num_events))`. Perhaps we could keep total use days data (combines alcohol, cannabis etc) and fill with 0?

```{r}
y2 |> select_if(~ is.factor(.)) |> summary() |> kbl() |> kable_styling()
```

Nobody has tried illicit drugs, so this can be removed. 

```{r}
y2 = y2 |> mutate(
  tlfb_cal_scr_num_events = replace_na(tlfb_cal_scr_num_events, 0)
)

y2 = y2 |> select_if(~ sum(is.na(.))<0.2*nrow(y2))
y2 = y2 |> select(-c(eventname, illicit, site_id_l, rel_birth_id, interview_date, visit_type, gender_id))

names(y2) = c("src_subject_id", "tlfb_use_days", "weightcontrol_ksads", "witness_comm_violence", "death_loved_one", "witness_dv", "s_abuse", "p_abuse", "emot_abuse", "serious_accident", "bkfs_fruit", "bkfs_veg", "bkfs_fiber", "sleep_hrs", "bmi", "needed_food", "income", "parent_ed", "birthsex", "gender", "race_ethnicity", "area_depriv", "comm_safety", "discrimination", "days_active", "bullying_victim", "cyberbullying", "chronotype", "life_events", "fam_conflict", "p_monitoring", "eff_control", "p_depression", "agemths", "cbcl_internalising", "cbcl_dsm5_depress")

y2 = droplevels(y2)

# remove cbcl missing data
y2 = y2 %>% filter(!is.na(cbcl_internalising))
sapply(y2, function(x) sum(is.na(x))) |> kbl(col.names = c("", "N missing")) |> kable_styling() |> scroll_box()
```

Little's MCAR test for continuous variables:

```{r}
y2 |> select_if(~ is.numeric(.)) |> naniar::mcar_test() |> kbl() |> kable_styling()
```

```{r}
## split into train and test samples ##
index = caret::createDataPartition(y2$cbcl_internalising[which(is.na(y2$cbcl_internalising)==F)], 
                            p = 0.2, list = F, times = 1)
traindat = y2[-index,]
testdat = y2[index,]
```

## Descriptive statistics

### Full sample

```{r}
y2[,-1] |>  dfSummary(plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```


### Training data:

```{r}
traindat[,-1] |> dfSummary(plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```

### Test data:

```{r}
testdat[,-1] |> dfSummary(plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```


```{r eval=FALSE}
saveRDS(traindat, "G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/y2_traindat.rds")
saveRDS(testdat, "G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/y2_testdat.rds")
```

