---
title: "Combining depRS and PRS for binary MDD"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, results = "asis")
library(tidyverse)
library(kableExtra)
library(performance)
library(pROC)
library(summarytools)
define_keywords(title.dfSummary = "NA", ask = F)
pal = paletteer::paletteer_d("colorblindr::OkabeIto")

source("C:/Users/eilee/Desktop/PhD/Year 3/envRS/SCRIPTS/summarystats.R")

set.seed(2025)
```

This script contains supplemental analyses predicting **binary** lifetime MDD at Year 2 from depRS and MDD PRS from the 2025 PGC GWAS. PRS for EUR participants are estimated using **European-only sumstats**, PRS for AFR, EAS and AMR participants are estimated using the **diverse ancestry sumstats**. PRS for AMR participants were estimated with the EUR ancestry LD matrix. depRS was trained to predict **continuous** CBCL DSM5-oriented depression subscale scores - measured at Year 2 - from 23 environmental factors measured at baseline. depRS and PRS are mean centred and standardised for comparison of ORs.

For (a) self-reported and (b) parent-reported lifetime MDD, we compare the following models: 

Single predictor:\
(m1depRS) Y2 MDD \~ depRS\
(m1PRS) Y2 MDD \~ PRS\

(m2) Y2 MDD \~ depRS + PRS\
(m3) Y2 MDD \~ depRS*PRS


```{r}
mdd2025 = readRDS("G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/mdd25_prs.rds")

for (anc in 1:length(mdd2025)) {
  ancestry = names(mdd2025)[anc]
  val = rep(ancestry, times = nrow(mdd2025[[anc]]))
  mdd2025[[anc]]$ancestry = as.factor(val)
}

PRS = bind_rows(mdd2025$afr, mdd2025$eas, mdd2025$amr_diverse, mdd2025$eur_european)

depRS = readRDS("G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/depRS_long.rds") |>
  select(src_subject_id, lambda.min)

scores = merge(depRS, PRS, by.x = "src_subject_id", by.y = "IID") |> 
  mutate(across(where(is.numeric), scale)) |> 
  mutate(across(where(is.matrix), as.numeric))
names(scores)[2:3] = c("depRS", "PRS")
```

depRS was available for N=`r nrow(depRS)` participants.\
PRS was available for N=`r nrow(PRS)` participants.\

A total N=`r nrow(scores)` participants had complete depRS and PRS data. Analyses will be conducted in this sample.

```{r}
rm(depRS, PRS)
```

# (a) Youth self-reported MDD

```{r}
mddy = readRDS("G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/ksads_y.rds")

dat.y = merge(scores, mddy$y2, by = "src_subject_id")

dat.y |> summarise(
  N = n(),
  LifetimeMDD = sum(mdd_lifetime=="1"),
  NoMDD = sum(mdd_lifetime=="0")
) |> kbl(digits = 3)

prs_youth = dat.y |> select(src_subject_id, mdd_lifetime, ancestry)
```

Of the N=`r nrow(dat.y)` participants with self-reported MDD data at year 2, there were N=`r nrow(filter(dat.y, mdd_lifetime==1))` lifetime MDD cases. 

## Model comparisons {.tabset}

```{r}
m1depRS.y = glm(mdd_lifetime ~ depRS, dat.y, family = binomial(link = "logit"))
m1PRS.y = glm(mdd_lifetime ~ PRS + PC1_AVG + PC2_AVG + PC3_AVG + PC4_AVG + PC5_AVG, dat.y, family = binomial(link = "logit"))

m2.y = glm(mdd_lifetime ~ depRS + PRS + PC1_AVG + PC2_AVG + PC3_AVG + PC4_AVG + PC5_AVG, 
         dat.y, family = binomial(link = "logit"))
m3.y = glm(mdd_lifetime ~ depRS*PRS + PC1_AVG + PC2_AVG + PC3_AVG + PC4_AVG + PC5_AVG,
         dat.y, family = binomial(link = "logit"))
```

### ~ depRS

```{r}
test_performance(m1depRS.y, m2.y, m3.y) |> display(digits=3)
```

### ~ PRS

```{r}
test_performance(m1PRS.y, m2.y, m3.y) |> display(digits=3)
```

## Performance metrics

```{r}
mods.y = list(m1depRS.y, m1PRS.y, m2.y, m3.y)

compare_performance(mods.y, verbose = F) |> 
  mutate(Name = c("depRS", "PRS", "depRS+PRS", "depRS*PRS")) |> 
  kable(digits = 3) |> kable_styling()

lapply(mods.y, performance_accuracy) |> 
  lapply(as.data.frame) |> 
  bind_rows(.id = "model") |> 
  mutate(model = c("depRS", "PRS", "depRS+PRS", "depRS*PRS")) |> 
  kable(digits = 3) |> kable_styling()
```

## Model estimates

```{r}
sjPlot::tab_model(mods.y, dv.labels = c("depRS", "PRS", "depRS+PRS", "depRS*PRS"), collapse.ci = T, show.se = T, show.stat = T, terms = c("depRS", "PRS", "depRS:PRS"))
```

## ROC curves

```{r}
#ROC curves for each model
dat.y = dat.y |> mutate(
  m1depRS = m1depRS.y$fitted.values,
  m1PRS = m1PRS.y$fitted.values,
  m2 = m2.y$fitted.values,
  m3 = m3.y$fitted.values
)

mods = c("m1depRS", "m1PRS", "m2", "m3")

rocs.y = list()
for (m in 1:length(mods.y)) {
  out = roc(dat.y$mdd_lifetime, dat.y[,mods[m]], 
            ci = TRUE, auc = TRUE)
  rocs.y[[mods[m]]] = out
}
```

```{r}
youth = ggroc(list("depRS" = rocs.y$m1depRS, "PRS" = rocs.y$m1PRS, "depRS + PRS" = rocs.y$m2, "depRS*PRS" = rocs.y$m3)) +
  scale_color_manual(values = pal) +
  geom_segment(aes(x=1, y=0, xend=0, yend=1), colour = "black", alpha = 0.7, linetype = "dashed") +
  theme_bw() + 
  labs(x = "Specificity", y = "Sensitivity", color = "Model")

youth

#par("din")
```

# (b) Parent-reported MDD

```{r}
mddp = readRDS("G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/ksads_p.rds")

dat.p = merge(scores, mddp$y2, by = "src_subject_id")

dat.p |> summarise(
  N = n(),
  LifetimeMDD = sum(mdd_lifetime=="1"),
  NoMDD = sum(mdd_lifetime=="0")
) |> kbl(digits = 3)

prs_parent = dat.p |> select(src_subject_id, mdd_lifetime, ancestry)
```

Of the N=`r nrow(dat.p)` participants with self-reported MDD data at year 2, there were N=`r nrow(filter(dat.p, mdd_lifetime==1))` lifetime MDD cases.

## Model comparisons {.tabset}

```{r}
m1depRS.p = glm(mdd_lifetime ~ depRS, dat.p, family = binomial(link = "logit"))
m1PRS.p = glm(mdd_lifetime ~ PRS + PC1_AVG + PC2_AVG + PC3_AVG + PC4_AVG + PC5_AVG, dat.p, family = binomial(link = "logit"))

m2.p = glm(mdd_lifetime ~ depRS + PRS + PC1_AVG + PC2_AVG + PC3_AVG + PC4_AVG + PC5_AVG, 
         dat.p, family = binomial(link = "logit"))
m3.p = glm(mdd_lifetime ~ depRS*PRS + PC1_AVG + PC2_AVG + PC3_AVG + PC4_AVG + PC5_AVG,
         dat.p, family = binomial(link = "logit"))
```

### ~ depRS

```{r}
test_performance(m1depRS.p, m2.p, m3.p) |> display(digits=3)
```

### ~ PRS

```{r}
test_performance(m1PRS.p, m2.p, m3.p) |> display(digits=3)
```

## Performance metrics

```{r}
mods.p = list(m1depRS.p, m1PRS.p, m2.p, m3.p)

compare_performance(mods.p, verbose = F) |> 
  mutate(Name = c("depRS", "PRS", "depRS+PRS", "depRS*PRS")) |> 
  kable(digits = 3) |> kable_styling()

lapply(mods.p, performance_accuracy) |> 
  lapply(as.data.frame) |> 
  bind_rows(.id = "model") |> 
  mutate(model = c("depRS", "PRS", "depRS+PRS", "depRS*PRS")) |> 
  kable(digits = 3) |> kable_styling()
```

## Model estimates

```{r}
sjPlot::tab_model(mods.p, dv.labels = c("depRS", "PRS", "depRS+PRS", "depRS*PRS"), collapse.ci = T, show.se = T, show.stat = T, terms = c("depRS", "PRS", "depRS:PRS"))
```

## ROC curves

```{r}
#ROC curves for each model
dat.p = dat.p |> mutate(
  m1depRS = m1depRS.p$fitted.values,
  m1PRS = m1PRS.p$fitted.values,
  m2 = m2.p$fitted.values,
  m3 = m3.p$fitted.values
)

mods = c("m1depRS", "m1PRS", "m2", "m3")

rocs.p = list()
for (m in 1:length(mods)) {
  out = roc(dat.p$mdd_lifetime, dat.p[,mods[m]], 
            ci = TRUE, auc = TRUE)
  rocs.p[[mods[m]]] = out
}
```


```{r}
parent = ggroc(list("depRS" = rocs.p$m1depRS, "PRS" = rocs.p$m1PRS, "depRS + PRS" = rocs.p$m2, "depRS*PRS" = rocs.p$m3)) +
  scale_color_manual(values = pal) +
  geom_segment(aes(x=1, y=0, xend=0, yend=1), colour = "black", alpha = 0.7, linetype = "dashed") +
  theme_bw() + 
  labs(x = "Specificity", y = "Sensitivity", color = "Model")

parent
```

# ROC figure

```{r}
library(patchwork)

ptch_plot = youth + theme(legend.position = "none") + ggtitle("(a)") | parent + ggtitle("(b)")
```

```{r}
# ggsave("C://Users/eilee/Desktop/PhD/Year 3/Drafts/envRS_paper/Figures/ROCs_sens.png", ptch_plot, device = "png", create.dir = TRUE, width = 9, height = 4, units = "in")
```

# Descriptives

## (a) Youth self-reported MDD {.tabset}

```{r}
y.dat = readRDS("G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/predictors_unrelatedIDs.rds") |> filter(src_subject_id %in% prs_youth$src_subject_id) |> select(src_subject_id, eventname, interview_age, gender_id, birthsex, race_ethnicity, income, parent_ed) |> mutate(income = as.factor(income))

y2age = y.dat |> filter(eventname == "2_year_follow_up_y_arm_1") |> mutate(y2_age = interview_age/12) |> select(src_subject_id, y2_age)
y.base = y.dat |> filter(eventname=="baseline_year_1_arm_1") |> mutate(age_yrs = interview_age/12)

prs_youth = merge(prs_youth, y2age)
prs_youth = merge(prs_youth, y.base)
```

```{r}
sumcols = c("age_yrs", "y2_age", "gender_id", "birthsex", "race_ethnicity", "income", "parent_ed", "ancestry", "mdd_lifetime")
all_y_sum = get_sum_stats(df = prs_youth, vars = sumcols)
case_y_sum = prs_youth |> filter(mdd_lifetime==0) |> get_sum_stats( vars = sumcols)
ctrl_y_sum = prs_youth |> filter(mdd_lifetime==0) |> get_sum_stats( vars = sumcols)
```


### Full sample

```{r}
summarytools::dfSummary(prs_youth[,-1],plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```

### Lifetime MDD \= 1

```{r}
prs_youth[,-1] |> filter(mdd_lifetime==1) |> 
summarytools::dfSummary(plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```

### Lifetime MDD \= 0

```{r}
prs_youth[,-1] |> filter(mdd_lifetime==0) |> 
summarytools::dfSummary(plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```


## (b) Parent-reported MDD {.tabset}

```{r}
p.dat = readRDS("G://users/eileen/ABCD/ABCD_Environmental_Risk/ABCDv5.1/DATA/predictors_unrelatedIDs.rds") |> filter(src_subject_id %in% prs_parent$src_subject_id) |> select(src_subject_id, eventname, interview_age, gender_id, birthsex, race_ethnicity, income, parent_ed) |> mutate(income = as.factor(income))

y2age = p.dat |> filter(eventname == "2_year_follow_up_y_arm_1") |> mutate(y2_age = interview_age/12) |> select(src_subject_id, y2_age)
p.base = p.dat |> filter(eventname=="baseline_year_1_arm_1") |> mutate(age_yrs = interview_age/12)

prs_parent = merge(prs_parent, y2age)
prs_parent = merge(prs_parent, p.base)
```

```{r}
sumcols = c("age_yrs", "y2_age", "gender_id", "birthsex", "race_ethnicity", "income", "parent_ed", "ancestry", "mdd_lifetime")
all_p_sum = get_sum_stats(df = prs_parent, vars = sumcols)
case_p_sum = prs_parent |> filter(mdd_lifetime==1) |> get_sum_stats( vars = sumcols)
ctrl_p_sum = prs_parent |> filter(mdd_lifetime==0) |> get_sum_stats( vars = sumcols)
```

```{r}
descriptives = list("all_youth" = all_y_sum, "case_youth" = case_y_sum, "ctrl_youth" = ctrl_y_sum, "all_parent" = all_p_sum, "case_parent" = case_p_sum, "ctrl_parent" = ctrl_p_sum)

openxlsx::write.xlsx(descriptives, "C://Users/eilee/Desktop/PhD/Year 3/Drafts/envRS_paper/sens_MDD_PRS_DESC.xlsx")
```

### Full sample

```{r}
summarytools::dfSummary(prs_parent[,-1],plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```

### Lifetime MDD \= 1

```{r}
prs_parent[,-1] |> filter(mdd_lifetime==1) |> 
summarytools::dfSummary(plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```

### Lifetime MDD \= 0

```{r}
prs_parent[,-1] |> filter(mdd_lifetime==0) |> 
summarytools::dfSummary(plain.ascii = F, style = "grid", valid.col = F, graph.col = T, varnumbers = F, tmp.img.dir = "/tmp", max.tbl.height = 300) |> print(method = "render")
```